
\begin{abstract}
激光与物质的相互作用是高能物理的很重要的分支，是新能源以及高能物理进一步发展很重要的方向。例如：受控核聚变，很有可能解决未来很长一段时间内的能源问题。激光加速器具有体积小造价低的优势，其应用涉及癌症治疗，新型辐射源以及成像技术等。随着强激光技术的逐渐成熟，这一领域步入飞速发展的阶段，趋于成熟的电子加速器，曾出不穷的辐射源方案，以及不断地突破的离子加速能量记录。在激光加速领域中，离子加速的提升空间还很大，束流品质难以有效控制。
本文研究强相对论激光与物质相互作用，以高性能计算为工具，围绕临界密度等离子体对激光离子加速的影响展开，依据现有实验室条件为参考，着重于离子能量提高以及束流品质改善，并对其应用进行讨论。论文主要分为三部分，可控式固体靶激光烧蚀，临界密度等离子体增加激光离子加速，强激光等离子体中的流不稳定性，以及激光加速质子在成像技术中的应用。

第一部分研究固体靶在激光脉冲作用下的可控烧蚀。金属固体靶在非相对论激光作用下，离化加热并膨胀，是激光与物质作用普遍的现象，在实验研究中有着重要的影响，其物理过程可由流体力学方法描述。通过研究激光以及固体靶参数对于烧蚀过程的影响，使得其密度分布由激光的光强以及脉冲周期控制，为后续的离子加速作为基础。流体力学模拟仿真程序MULTI是重要工具。
 
第二部分提出了用临界密度等离子体控制激光加速的方法。利用第一部分中金属靶烧蚀，控制临界密度等离子体的生成，提高激光与等离子体的能量耦合，最终实现对于离子加速的增强作用。理论分析以及PIC 计算机模拟相得益彰。

第三部分分析了在强激光产生的相对论电子在等离子中传播时的流不稳定性以及对于质子加速的影响。强激光与物质作用过程中，不稳定性是普遍存在的，一定条件下，其增长速度急剧提高，对于激光与等离子体的能量耦合以及质子加速有着重要的影响。
第四部分整合前基本分的结果，强调了应用。

\keywords{激光，等离子体，离子加速，不稳定性，}
\end{abstract}


\begin{englishabstract}
Laser matter interaction is a hot topic for the new energy sources and new accelerator, for example: Inertial Confined Fusion is the potential candidates for the safty energy source and laser accelerators have the advantages in size and cost, application in cancer therapy, radiation source and radiography technology and so on.  As the developing of the laser technology, it's now fast-growing. The mature laser electron accelerator, new radiation source scheme and upgrading ion energy record. However we have still large room improving the ion energy, emittance etc and control them.

Thsis thesis investigate the ultra-intense laser interaction with matter by large massive computation,  focusing on the critical plasma effect on the ion acceleration enhancement and aiming for ion energy and other beam quality parameters improvement. Based on the nowadays available experiment condition, it includes also the application. The paper contains three parts: controllable laser foil target ablation; laser ion acceleration enhancement by critical plasma; The current instability during the laser plasma interaction and the effect on ion acceleration; The applicaiton in proton radiography.

The first section research the metal foil target ablation by laser pulse. In non-relativistic laser regime, it's common to have foil target ionization, heating and expansion. And it's an important part in experiment. Hydrodynamic model is effective to analysis the process, MULTI code is the tool. We get the relationship between target expansion with laser and target parameters. It can be used to control the target expansion for the future acceleration.


The second section propose the scheme to control ion acceleration by critical plasma. By controlling the target ablation discribed in the obove section, it's feasible to get the most effective laser plasma energy coupling, and finally benifit to the ion acceleration. The theoretical analysis and the  PIC simulation agree to each other.

The third section we investigate the current instability during the process that laser driven current  propagates in plasma. and the effect on ion acceleration. Current instability is universal during the laser plasma interaction. The growing can be dramatic under specific conditon, effecting much on ion acceleration. 

The fourth section is for application. Using the ion beam got from the above scheme for proton radiography and so on. 


 

\englishkeywords{laser, plasma, ion acceleration, instability}
\end{englishabstract}
